---
title: "Amino acid position plots"
author: "mfpfox"
date: "4/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GOAL: explore positional differences for proteins using line plots

## columns: 

### notes: 
- cadd scores for GRCh38 come from dbNSFP accession fltered files
- cadd version v1.4 both models
- dbnsfp version 4.0a
- LIST file provided conservation and disorder score
- only missense CADD scores (filtered out nonsense)
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(RColorBrewer)
library(plotly)
library(tidyr)
```

```{r}
df <- read_csv('G6PDfigure_CADDmaxmean_7CKcodonclean.csv')
df
```

```{r}
summary(df)
```

# add column for cadd max-mean, both GRCh37 and GRCh38
```{r}

```

## what has highest difference (CADD MAX - CADD MEAN)
```{r}
df <- df %>% 
  filter(aapos_id != 'X_516')
df %>% 
  filter(max_minus_avg > 8)
```

### split df
```{r}
df <- df %>%
  separate(aapos_id, c('aaref','aapos'), sep = "_")
```

```{r}
df[order(df$cadd38_max_missense,decreasing = TRUE),c(1,2,3,4,5,6)]
```

## fix column type 
```{r}
df$aapos <- as.numeric(df$aapos)
```



# max and average
```{r}
lineaa <- df %>%
  ggplot(aes(x=aapos, y=cadd38_max_missense)) + 
  theme_minimal() + theme(legend.position = "none") + theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +  ylab("CADD38 Scores") +  xlab("G6PD AA position") + 
  theme(plot.title = element_text(hjust=0.5), axis.title.x = element_text(size=13, color="black", margin=margin(t=15, b=5)),axis.title.y = element_text(size=11, color="black", margin=margin(t=0, r=15, b=0)), axis.text=element_text(size=11, color="black",margin=margin(t=15, b=5))) + 
  geom_line(data = df, aes(x = aapos, y = cadd38_max_missense), size=.4,color = "darkslategray3") +
  geom_line(data = df, aes(x = aapos, y = cadd38_avg_missense), size=.4,color = "darkslateblue") +
  geom_vline(xintercept=c(497,446,432,408,386,385,294,205,171,158,97,89,82,47,13), color="darkred", size=.2, alpha=0.9)  + scale_x_continuous(breaks=seq(0, 515, 50)) + scale_y_continuous(breaks=seq(0, 35, 5)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

lineaa
#ggsave("G6PDline_max_avg.pdf", width = 10, height=3, dpi= 300)

```

# average only
```{r}
lineavg <- df %>%
  ggplot(aes(x=aapos, y=cadd38_max_missense)) + theme_minimal() + theme(legend.position = "none") + theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  ylab("CADD38 average scores") +
  xlab("G6PD AA position") + theme(plot.title = element_text(hjust=0.5), axis.title.x = element_text(size=13, color="black", margin=margin(t=15, b=5)),axis.title.y = element_text(size=11, color="black", margin=margin(t=0, r=15, b=0)), axis.text=element_text(size=11, color="black",margin=margin(t=15, b=5))) + 
  geom_line(data = df, aes(x = aapos, y = cadd38_avg_missense), size=.4,color = "darkslateblue") +
  geom_vline(xintercept=c(497,446,432,408,386,385,294,205,171,158,97,89,82,47,13), color="darkred", size=.2, alpha=0.9)  + scale_x_continuous(breaks=seq(0, 515, 50)) + scale_y_continuous(breaks=seq(0, 35, 5)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

lineavg
#ggsave("G6PDline_avg_cadd38.pdf", width = 10, height=3, dpi= 300)
```

# max only 

```{r}
linemax <- df %>%
  ggplot(aes(x=aapos, y=cadd38_max_missense)) + theme_minimal() + theme(legend.position = "none") + theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  ylab("CADD38 max scores") +
  xlab("G6PD AA position") + theme(plot.title = element_text(hjust=0.5), axis.title.x = element_text(size=13, color="black", margin=margin(t=15, b=5)),axis.title.y = element_text(size=11, color="black", margin=margin(t=0, r=15, b=0)), axis.text=element_text(size=11, color="black",margin=margin(t=15, b=5))) + 
  geom_line(data = df, aes(x = aapos, y = cadd38_max_missense), size=.4,color = "darkslategray3") +
  geom_vline(xintercept=c(497,446,432,408,386,385,294,205,171,158,97,89,82,47,13), color="darkred", size=.2, alpha=0.9)  + scale_x_continuous(breaks=seq(0, 515, 50)) + scale_y_continuous(breaks=seq(0, 35, 5)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

linemax
#ggsave("G6PDline_max_cadd38.pdf", width = 10, height=3, dpi= 300)
```


```{r}
maxavg <- df %>%
  ggplot(aes(x=aapos, y=max_minus_avg)) + theme_minimal() + theme(legend.position = "none") + theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  ylab("CADD max - average") +
  xlab("G6PD AA position") + theme(plot.title = element_text(hjust=0.5), axis.title.x = element_text(size=13, color="black", margin=margin(t=15, b=5)),axis.title.y = element_text(size=11, color="black", margin=margin(t=0, r=15, b=0)), axis.text=element_text(size=11, color="black",margin=margin(t=15, b=5))) + 
  geom_line(data = df, aes(x = aapos, y = max_minus_avg), size=.4,color = "plum") +
  geom_vline(xintercept=c(497,446,432,408,386,385,294,205,171,158,97,89,82,47,13), color="darkred", size=.2, alpha=0.9)  + scale_x_continuous(breaks=seq(0, 515, 50)) + scale_y_continuous(breaks=seq(0, 9, 2)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

maxavg
ggsave("G6PDline_max_minus_avg.pdf", width = 10, height=2, dpi= 300)

```


# box plot of every letter as the X axis and max cadd scores are y axis

```{r}
boxaa <- df %>%
  mutate(aaref = fct_reorder(aaref, cadd38_max_missense, .fun='median')) %>%
  ggplot(aes(x=aaref, y=cadd38_max_missense, fill=aaref, alpha=0.5)) + theme_minimal() + theme(legend.position = "none") + theme(panel.grid.minor = element_blank(),panel.grid.major = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  ylab("Max CADD38 score per AA") +
  xlab("G6PD residues lost") + theme(plot.title = element_text(hjust=0.5), axis.title.x = element_text(size=15, color="black", margin=margin(t=15, b=5)),axis.title.y = element_text(size=15, color="black", margin=margin(t=0, r=15, b=0)), axis.text=element_text(size=12, color="black",margin=margin(t=15, b=5))) +  scale_y_continuous(breaks=seq(0, 40, 5)) +
  geom_boxplot(
  outlier.colour = "red",
  outlier.fill="red",
  outlier.size = 0.2) 
  
boxaa
#ggsave("G6PDbox_max_cadd38.pdf", width = 5, height=4, dpi= 300)
```

